---
title: "Week 9 data 607"
author: "Nikoleta Emanouilidi"
date: "2024-03-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the desired libraries.

```{r}
library(httr)
library(jsonlite)
library(tidyverse)

```

### The code below connects to the API.It takes the URL that is used for accessing data from the New York Times Most Popular API and sets a variable called "URL". 

```{r}
api_key <- "vdXCYZ9JGtiidjpx4FNEcOc5GHVnaJ5l"

URL <- paste0("https://api.nytimes.com/svc/mostpopular/v2/viewed/", 7, ".json") 
query_params <- list("api-key" = api_key)
request_url <- modify_url(URL, query = query_params) ### sends the request

response <- GET(request_url) 
response$status_code ### checks the response
```

### A response code of 200 means "OK" and it indicates that the request has succeeded.

```{r}
Articles_untidy <- response$content %>%
  rawToChar() %>%
  fromJSON(flatten = TRUE) %>%
  as.data.frame()
head(Articles_untidy)

```

### The code above fetches data from the server, converts it from raw format to character format, parses it as JSON data and converts it into a data frame.

```{r}
Articles_tidy <- Articles_untidy %>% 
  select(
    "Published Date" = results.published_date,
    "Popular Articles" = results.title,
    "Section" = results.section,
    "Abstract" = results.abstract
  )
View(Articles_tidy)
```

###  The code above takes the Articles_untidy dataset, selects specific columns and renames them.