---
title: "Week 10 Data 607"
author: "Nikoleta Emanouilidi"
date: "2024-03-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### *Load the libraries*
```{r}
### I found the libraries by reading chapter 1
library(tidyverse)
library(tidytext)
library(dplyr)
library(janeaustenr)
library(stringr)
library(ggplot2)
library(wordcloud)
library(reshape2)
```

### **Get sentiments datasets**
```{r}
## AFINN
get_sentiments("afinn")
```
```{r}
### BING
get_sentiments("bing")
```
```{r}
### NRC
get_sentiments("nrc")
```

### *Sentiment analysis with inner join*
```{r}
### Tidy the data
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

##sentiment analysis

nrc_joy <- get_sentiments("nrc") %>%
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)

```
### *negative and positive sentiments*
```{r}
jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
### Plot
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

### *Comparing the three sentiment dictionaries*

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")
```
```{r}
bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)

```
### Plot
```{r}
bind_rows(
  afinn,
  bing_and_nrc
) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")

```

```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)

get_sentiments("bing") %>% 
  count(sentiment)
```

### *Most common positive and negative words*
```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()
bing_word_counts

bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

### *stop-words*
```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

```
### *wordcloud*
```{r}
tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

```
```{r}

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)

```

### *Looking at units beyond just words*
```{r}
p_and_p_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")
p_and_p_sentences$sentence[2]

```

### Split the text of Jane Austen’s novels into a data frame by chapter.
```{r}
austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())
```

###  Number of negative words in each chapter and divide by the total words in each chapter
```{r}
bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  slice_max(ratio, n = 1) %>% 
  ungroup()
```

##  **Code Extension**
### I found an additional sentiment lexicon in https://bookdown.org/psonkin18/berkshire/sentiment.html (chapter 4.7) called **loughran**.I am using the hotels_review dataset from https://data.world/datafiniti/hotel-reviews. The file was more than 100 Mb and I couldn't upload it on github, so I removed some of the columns and uploaded the modified file.
```{r}
hotels <- read.csv("https://raw.githubusercontent.com/NikoletaEm/607LABS/main/modified_hotels_reviews.csv")
reviews<- hotels[,c(12,21)]
```
#### Let's start the sentiment analysis using **loughran**
```{r}
loughran_lexicon <- get_sentiments("loughran")

tokenized_reviews <- reviews %>%   ### we tokenize the reviews
  unnest_tokens(word, reviews.title)


sentiment_scores <- tokenized_reviews %>% ### we join with the sentiment lexicon to get sentiment scores
  inner_join(loughran_lexicon)

head(sentiment_scores)
```

### The code below calculates the difference between the count of positive and negative sentiment words, giving us a sentiment score. Positive scores indicate a positive sentiment, negative scores indicate a negative sentiment and zero indicates a neutral sentiment.
```{r}
sentiment_analysis <- sentiment_scores %>%
  group_by(name) %>%
  summarise(sentiment_score = sum(sentiment == "positive") - sum(sentiment == "negative"))

head(sentiment_analysis)
```

```{r}
review_counts <- sentiment_scores %>%
  count(sentiment)

ggplot(review_counts, aes(x = sentiment, y = n, fill = sentiment)) +
  geom_bar(stat = "identity", width = 0.5) +
  labs(title = "Distribution of Positive and Negative Reviews",
       x = "Sentiment",
       y = "Count") +
  scale_fill_manual(values = c("positive" = "skyblue", "negative" = "salmon")) +
  theme_minimal()
```

### As we see from the plot the positive reviews outnumber the negative ones , which means that the majority of the hotels was good.

```{r}
best_hotel <- sentiment_analysis[which.max(sentiment_analysis$sentiment_score), ]
best_hotel
```
### The hotel with the **most positive points** (highest sentiment_score) turns out to be Hyatt House Seattle/Downtown!